---
title: "Flux analysis"
author: "Luyao Kang"
date: "`r Sys.Date()`"
output:
  html_notebook:
    toc: yes
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include = F}
knitr::opts_chunk$set(
  collapse = T, echo = T, comment = "#>", message = F, warning = F,
	fig.align = "center", fig.width= 4.5, fig.height = 3, dpi = 300)
```

## Data input
Set work directory
```{r}
setwd('e:/thermokarst_gully/')
save.dir <- file.path(getwd(),"result")
```

Loading packages
```{r}
pacman::p_load(tidyverse, lme4)
```

Data input
```{r}
source("script/read_data_rdp.R")
```

Two way repeated measures analysis to test the difference in flux for the whole growing season. First is CO<sub>2</sub>:
```{r}
CO2_df <- flux_tab %>% filter(Flux_type == "CO2") %>%
  select(c(Time, Gully_id, Rep, Control, Collapsed)) %>%
  pivot_longer(cols = c("Control", "Collapsed"), names_to = "Group", values_to = "Flux_value")

stress.aov <- with(CO2_df,
                   aov(Flux_value ~ Group * Gully_id +
                         Error(Time / (Group * Gully_id)))
)
summary(stress.aov)
```
CH<sub>4</sub>:
```{r}
CH4_df <- flux_tab %>% filter(Flux_type == "CH4") %>%
  select(c(Time, Gully_id, Rep, Control, Collapsed)) %>%
  pivot_longer(cols = c("Control", "Collapsed"), names_to = "Group", values_to = "Flux_value")

stress.aov <- with(CH4_df,
                   aov(Flux_value ~ Group * Gully_id +
                         Error(Time / (Group * Gully_id)))
)
summary(stress.aov)
```
N<sub>2</sub>O:
```{r}
N2O_df <- flux_tab %>% filter(Flux_type == "N2O") %>%
  select(c(Time, Gully_id, Rep, Control, Collapsed)) %>%
  pivot_longer(cols = c("Control", "Collapsed"), names_to = "Group", values_to = "Flux_value")

stress.aov <- with(N2O_df,
                   aov(Flux_value ~ Group * Gully_id +
                         Error(Time / (Group * Gully_id)))
)
summary(stress.aov)
# using lme within nlme package
library(nlme)
lme_fitn2o.block <- lme(fixed = Flux_value ~ Group * Time ,  
                        random = ~1|Rep/Gully_id, 
                        method='REML',
                        corr = corAR1(form= ~1|Rep/Gully_id),
                        data = N2O_df)
summary(lme_fitn2o.block)
car::Anova(lme_fitn2o.block, type="III")
```

Determine the mean value of flux for the whole growing season.
```{r}
flux_ave_df <- flux_tab %>% select(c("Flux_type", "Gully_id", "Rep", "Control", "Collapsed")) %>%
  group_by(Flux_type, Gully_id, Rep) %>%
  dplyr::summarise(across(, mean, na.rm = TRUE))
```

Test the difference in flux between control and collapsed using linear mixed model. First is CO<sub>2</sub>:
```{r}
library(lme4)
library(car)
CO2_df <- flux_ave_df %>% filter(Flux_type == "CO2") %>%
  pivot_longer(cols = c("Control", "Collapsed"), names_to = "Group", values_to = "Flux_value")
fm1 <- lmer(Flux_value ~ Group + (1 | Gully_id), data = CO2_df)
summary(fm1)
car::Anova(fm1, type = 2)
```
Then, test the difference in CH<sub>4</sub>.
```{r}
library(lme4)
library(car)
CH4_df <- flux_ave_df %>% filter(Flux_type == "CH4") %>%
  pivot_longer(cols = c("Control", "Collapsed"), names_to = "Group", values_to = "Flux_value")
fm2 <- lmer(Flux_value ~ Group + (1 | Gully_id), data = CH4_df)
summary(fm2)
car::Anova(fm2, type = 2)
```
Finally, test the difference in N<sub>2</sub>O.
```{r}
library(lme4)
library(car)
N2O_df <- flux_ave_df %>% filter(Flux_type == "N2O") %>%
  pivot_longer(cols = c("Control", "Collapsed"), names_to = "Group", values_to = "Flux_value")
fm3 <- lmer(Flux_value ~ Group + (1 | Gully_id), data = N2O_df)
summary(fm3)
car::Anova(fm3, type = 2)
```

```{r, fig.align='center', fig.width=7.5, fig.height=3.5}
#load the packages
pacman::p_load(ggplot2, ggpubr, gghalves)
# set the main theme
main_theme = theme_bw() + 
  theme(panel.grid = element_blank(),
        panel.border = element_rect(size = 0.5),
        strip.text = element_text(colour = 'black', size = 6),
        strip.background = element_rect(colour = 'black', fill = 'grey'),
        axis.title = element_text(color = 'black',size = 6),
        axis.ticks = element_line(color = "black", linewidth = 0.5),
        axis.text.y = element_text(colour = 'black', size = 6),
        axis.text.x = element_text(colour = 'black', size = 6),
        legend.position = "none")
# set the contrast group for wilcox.test
my_comparisons <- list(c('Control', 'Collapsed'))
flux_names <- list(
  'CO2' = expression(paste("CO"[2], " (", mu, "mol", " m"^-2, " s"^-1, ")")),
  'CH4'= expression(paste("CH"[4], " (nmol", " m"^-2, " s"^-1, ")")),
  'N2O'= expression(paste("N"[2], "O", " (nmol", " m"^-2, " s"^-1, ")"))
)

# write a function to change the facet labels
facet_labeller <- function(variable,value){
  return(flux_names[value])
}

# plot
flux_plot <- flux_ave_df %>%
  select(c("Flux_type", "Gully_id", "Rep", "Control", "Collapsed")) %>%
  pivot_longer(cols = c("Control", "Collapsed"), names_to = "Group", values_to = "Flux_value") %>% 
  mutate(Flux_type = factor(Flux_type, levels = c("CO2", "CH4", "N2O"))) %>%
  mutate(Group = factor(Group, levels = c("Control", "Collapsed"))) %>%
  ggplot(aes(Group, Flux_value, fill = Group)) +
  geom_half_violin(position = position_nudge(x = 0.25), side = "r", width = 0.8, color = NA) +
  geom_boxplot(width = 0.4, size = 0.75, outlier.color = NA) +
  geom_jitter(aes(fill = Group), shape = 21, size = 1.5, width = 0.2) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  stat_compare_means(comparisons = my_comparisons, paired = F,
                     p.adjust.method = "BH", label = "p.signif", bracket.size = 0.3,
                     size = 3.5, tip.length = 0.00, method = "wilcox.test") +
  labs(x = NULL, y = NULL) +
  scale_fill_manual(values = c("#79ceb8", "#e95f5c", "#5cc3e8", "#ffdb00")) +
  facet_wrap(~Flux_type, scales = "free_y", ncol = 3, labeller = facet_labeller) +
  main_theme
if (!dir.exists(file.path(save.dir, "figs/flux/"))) {
  dir.create(file.path(save.dir, "figs/flux/"))
}
# ggsave(file.path(save.dir, "./figs/flux/flux_plot.pdf"),
#        flux_plot, width = 4.5, height = 2, units = "in")
flux_plot
```

### test the difference in abundance of main taxa using liner models
```{r}
# determine the effect size of the permafrost thawing on each phylum
env_scale <- metadata %>% 
  select(-c("Sample_id", "Sample_name")) %>%
  mutate(across(where(is.numeric), scale)) %>%
  mutate(Group = factor(Group, levels = c("Control", "Collapsed"))) %>%
  select(where(~ !any(is.na(.))))

# codes for calculating the effect size refer to wu et al. 2022:https://github.com/Linwei-Wu/warming_soil_biodiversity.
env_S1 <- sapply(3:ncol(env_scale), function(j) {
    message("Now j=", j, " in ", ncol(env_scale), ". ", date())
    if (length(unique(env_scale[, j])) < 3) {
        result <- rep(NA, 10)
    } else {
        fm1 <- lmer(env_scale[, j] ~ Group + (1 | Gully_id), data = env_scale)

        presult <- car::Anova(fm1, type = 2)
        coefs <- coef(summary(fm1))[, "Estimate"]  ##four coefs
        names(coefs) <- paste0(names(coefs), ".mean")

        SEvalues <- coef(summary(fm1))[, "Std. Error"]  ##standard errors
        names(SEvalues) <- paste0(names(SEvalues), ".se")

        tvalues <- coef(summary(fm1))[, "t value"]  ##t values
        names(tvalues) <- paste0(names(tvalues), ".t")

        chisqP <- c(presult[, 1], presult[, 3])
        names(chisqP) <- c(paste0(row.names(presult), ".chisq"), paste0(row.names(presult), ".P"))

        result <- c(coefs, tvalues, SEvalues, chisqP)
    }
    result
})
colnames(env_S1)<-colnames(env_scale)[-c(1:2)]
data.frame(env_S1)
```

Effect size plot
```{r, fig.align='center', fig.width=3.5, fig.height=4.5}
# rep_str = c("Plant_richness" = "Plant Richness",
#             "Plant_shannon" = "Plant Shannon",
#             "Soil_moisture" = "Soil moisture",
#             "H2O2" = expression(paste("H"[2], "O"[2])))
p.stars <- function(p.values) {
  unclass(symnum(p.values, corr = FALSE, 
                 na = FALSE, cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1),
                 symbols = c("***", "**", "*", ".", " ")))}
diff_in_env_plot <- env_S1 %>%
  t() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(., "variables") %>%
  mutate(sig = as.vector(unlist(lapply(Group.P, p.stars)))) %>%
  mutate(variables = factor(variables, levels = colnames(metadata)[-c(1:4)])) %>%
  ggplot(aes(x = variables, y = GroupCollapsed.mean, color = variables)) +
  geom_hline(aes(yintercept =0), size=0.7,  colour="gray2")+
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = GroupCollapsed.mean - GroupCollapsed.se, 
                    ymax = GroupCollapsed.mean + GroupCollapsed.se), 
                width = 0, position = position_dodge(width = 0.7), cex = 0.9) +
  geom_text(aes(label = sig, x = variables, y = (GroupCollapsed.mean/abs(GroupCollapsed.mean))*(abs(GroupCollapsed.mean) + GroupCollapsed.se)*1.2),
            position = position_dodge(0.1), vjust = 0.55) +
  labs(x = NULL, y = "Effect size")+
  # scale_color_manual(values=c("#D55E00","#446DA9","#446DA9","#446DA9","#446DA9","#D55E00","#D55E00")) +
  scale_y_continuous(expand = c(0, 0), limit = c(-2, 2)) +
  theme_bw() + coord_flip() + scale_x_discrete(position = "top") +
  annotate("rect", xmin = 0, xmax = 2.5, ymin = -2, ymax = 2, alpha = 0.2, fill = "#ff6666") +
  annotate("rect", xmin = 2.5, xmax = 3.5, ymin = -2, ymax = 2, alpha = 0.2, fill = "#f19837") +
  annotate("rect", xmin = 3.5, xmax = 8.5, ymin = -2, ymax = 2, alpha = 0.2, fill = "#e56eee") +
  annotate("rect", xmin = 8.5, xmax = 10.5, ymin = -2, ymax = 2, alpha = 0.2, fill = "#5fb236") +
  main_theme +
  theme(strip.background = element_rect(fill = c("#FFF6E1")))
if (!dir.exists(file.path(save.dir, "figs/env/"))) {
  dir.create(file.path(save.dir, "figs/env/"))
}
ggsave(file.path(save.dir, "/figs/env/diff_in_env_plot.pdf"),
       diff_in_env_plot, width = 3, height = 3.5, units = "in")
diff_in_env_plot
```

Combine plots
```{r, fig.align='center', fig.width=8.9, fig.height=15}
# box plot
my_comparisons <- list(c('Control', 'Collapsed'))
p4 <- metadata %>%
  select(c("Gully_id", "Group", "BG",  "NAG", "LAP", "PPO", "H2O2", "WHC", "SOC", "NO3_N", "NH4_N", "AP", "AGB", "plant_pathogen_control")) %>%
  pivot_longer(-c("Gully_id", "Group"), names_to = "functions", values_to = "values") %>%
  mutate(functions = factor(functions, levels = c("BG",  "NAG", "LAP", "PPO", "H2O2", "WHC", "SOC", "NO3_N", "NH4_N", "AP", "AGB", "plant_pathogen_control"))) %>%
  mutate(Group = factor(Group, levels = c("Control", "Collapsed"))) %>%
  mutate(Gully_id = factor(Gully_id, levels = c('EB', 'ML', 'RS', 'SLH', 'HSX', 'HH'))) %>%
  ggplot(aes(x = Group, y = values)) + 
  geom_boxplot(width = 0.5, aes(fill = Group), outlier.shape = NA) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15))) +
  stat_compare_means(comparisons = my_comparisons, paired = F,
                     p.adjust.method = "BH", label = "p.signif", bracket.size = 0.5,
                     size = 3.5, tip.length = 0.00, method = "wilcox.test") +
  labs(x = 'Group', y = NULL, fill= 'Group') +
  facet_grid(functions ~ Gully_id, scales = "free", space = "free_x") +
  scale_fill_manual(values = c("#79ceb8", "#e95f5c", "#5cc3e8", "#ffdb00")) +
  main_theme +
  theme(panel.spacing = unit(0, "lines"),
        axis.text.x = element_text(angle = 45, hjust = 1))
p4
```

